\input{../../doc}
\input{../../preamb}
\input{../eng}

\begin{document}
\opgt

\op{opggangutv}
Calculate by writing the numbers as sums of ones, tens and hundreds, and applying the distributive law.
\opgeks{ \vspace{5pt}
$ 15\cdot 3=(10+5)\cdot3=10\cdot3+5\cdot3 =30+15=45 $  \vsk

$ 147\cdot2=(100+40+7)\cdot 2= 100\cdot2 +40\cdot2+7\cdot2=200\cdot80+14=294 $
} \vspace{5pt}
\abch{
\item $ 17\cdot2 $
\item $ 59\cdot 3 $
\item $ 25\cdot 4 $ 
\item $ 582\cdot 2 $
\item $ 981\cdot3 $ 
}

\nes

\op{opgfaktto}
Write the numbers as a multiplication between two factors.\os
\abch{
	\item 100
	\item 30
	\item 40
	\item 70
} \os
\abchs{5}{
	\item 42
	\ \ \item 32
	\item 84
	\item 90
}

\begin{comment}
	\op{opgprimfakt}
	Primtalsfaktoriser. \os
	\abch{
	\item 36 
	\item 81
	\item 42
	\item 38
	\item 98
	\item 33 
	} \os
	\abchs{7}{
	\item 144 
	\item 129
	\item 198
	\item 555
	\item 888
	\item 930
	}
\end{comment}

\op{opgprimfakt}
Prime factorize the numbers from \refops{opgfaktto}. \\
\mers{It is recommended that the reader find it's own ways of prime factorization, but those who insist on following a schematic method can look up \refops{opgforklprimfakt}}.

\op{opgfakttre}
Factorize the numbers from \refops{opgfaktto} in three different ways.

\op{opgperf}
6 is called a \outl{perfect number} because the sum of it's factors 6 \\(including 1, but excluding 6) equals 6: $ 1+2+3=6 $. Find the next perfect number (it lies between 15 og 30).
\newpage
\op{opgforklprimfakt}
The examples below shows a method for prime factorization of numbers. Explain the method.
\begin{center}
	\parbox{0.35\linewidth}{
		\eks[1]{
			$ \colo{84}= \colb{2\cdot2\cdot3\cdot7}$	 \vsk
			
			\begin{tabular}{r|r|r}
				& $ : $&\\ \hline
				\colo{84}&\colb{2}&\colg{42} \\
				\colg{42}& \colb{2} &\colg{21} \\ 
				\colg{21}&\colb{3}&\colb{7}\\
				\hline
			\end{tabular}
	}} \qquad
	\parbox{0.35\linewidth}{
		\eks[2]{
			$ \colo{595}= \colb{5\cdot7\cdot17}$	 \vsk
			
			\begin{tabular}{r|r|r}
				& $ : $&\\ \hline
				\colo{595}& \colb{5} &\colg{119} \\ 
				\colg{119}&\colb{7}&\colb{17}\\
				\hline
			\end{tabular}
			\vspace{12pt}
	}}
\end{center} 


\newpage
\grubop{opgoddgangodd}
Explain why the product of to odd numbers always equals an odd number.

\grubop{opgfindprime}
Explain a method for checking if a number is a prime number.
\end{document}

